---
title: "Automatic vs Manual Transmission in term of MPG"
author: "Ameen AboDabash"
date: "Sep 2, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
##loading requrired packages
packages<-c("graphics","ggplot2","datasets")
sapply(packages, require, character.only = TRUE)
data(mtcars)
## setting the seed, to insure this analysis will end up with the same results
set.seed(1988)

# am as factor data
mtcars$am <- as.factor(mtcars$am)

# [,9]   am  Transmission (0 = automatic, 1 = manual)
levels(mtcars$am) <-c("Automatic", "Manual")

```

####Executive Summary
We'll explorer *mtcars* dataset to build a regression model which will confirm Manual is better than Automatic transitions in term of MPG "Miles Per Gallon" (avg 2.9 mpg much better ), and quantify that difference.

####Setting the Scene
Let's look at the data first, *mtcars* data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models), So  *mtcars* is A data frame with 32 observations on 11 variables  (Fig.1 Scatter plot for all vars)


[, 1] mpg *Miles/(US) gallon*       | [, 9] am *Transmission (0=automatic, 1=manual)*
------------------------------------|----------------------------------
[, 2] cyl  *Number of cylinders* | [, 3] disp *Displacement (cu.in.)*   
[, 4] hp *Gross horsepower*      | [, 5] drat *Rear axle ratio* 
[, 6] wt *Weight (lb/1000)*      | [, 7] qsec *1/4 mile time*
[, 8] vs *V/S*                   | [,10] gear *Number of forward gears*
[,11] carb *Number of carburetors* |  . 


Frankly speaking, its seems Manual transmission better than Automatic from data plotting "Appendix Fig.1", So lets get the *P-Value* to confirm this "we can call it" assumption or hypothesis:


```{r  }
t.test(mpg~am,mtcars,paired=FALSE,var.equal=FALSE)
```

So, P-Value is  0.001374 which s less than 0.05 which confirm our first assumption.


####Regression Model and Qunitifiying the difference
So the following basic model which multivariate linear regression with all variables:

```{r  }
summary(lm(mpg~am,mtcars))
 
```

Indeed, it looks like wt is the only factor that significantly changes with mpg. However, including all variables will possibly result over-fitting, R has an automatic variable-selection function *step*. so lets enhance our model using *step* as following:

```{r  }

summary(step(lm(data = mtcars, mpg ~ .), trace=0))
 
```
So *wt (weight)* included by step function and *qsec (1/4 mile time)* as long as *am (Transimession Type)* , *wt* negatively changes with *mpg*, and *qsec* and *am* positively changes.
Every lb/1000 weight increase will cause a decrease of roughly 4 mpg, every increase of 1/4 mile time will cause an increase of 1.2 mpg, and on average, manual transmission is *2.9 mpg better* than automatic transmission.Appendix(Fig.2).


####Appendix

######Fig.1
```{r fig.width=12, fig.height=5 }
require(graphics)
pairs(mtcars, main = "mtcars data")
#coplot(mpg ~ am | as.factor(am), data = mtcars, panel = panel.smooth, rows = 1)
boxplot(mpg~am, data = mtcars,
        xlab = "Transmission",
        ylab = "MPG",
        main = "MPG in term of MPG")

```


######Fig.2
```{r fig.width=12, fig.height=5 }
par(mfrow = c(2,2))
plot(step(lm(data = mtcars, mpg ~ .), trace=0))
```






Analysis available on [github]<https://github.com/aabodabash/autoVsManualMPG.git>
